---
title: "Final Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}

library(flexdashboard)
library(shiny)
library(leaflet)
library(sf)
library(readr)
library(dplyr)
library(tidyverse)
library(plotly)

#read pre-saved spatial RDS files
Obesity<- readRDS("obesity_geo.rds")
LPA <- readRDS("lpa_geo.rds")

```

# About the Data {data-navmenu="About"}

### About the Data

**Data Source:** [CDC PLACES Data Portal](https://data.cdc.gov/500-Cities-Places/PLACES-Local-Data-for-Better-Health-County-Data-20/swc5-untb/about_data)\
**Data Collection Method:** Estimates derived from the 2022 Behavioral Risk Factor Surveillance System (BRFSS), U.S. Census 2020 data, and 2018–2022 ACS using small area estimation techniques\
**Study Population:** U.S. adults across the 500 largest cities/counties\
**Time Period:** Data reflect responses from BRFSS 2021 and 2022, released in 2024

# Real-World Impact {data-navmenu="Real-World Impact"}
**Real-World Impact:**
This dashboard provides an accessible way to explore disparities in obesity and physical inactivity across U.S. counties, helping public health professionals identify high-need areas. The visualization of geographic patterns supports data-driven decision-making for targeted physical activity and obesity-focused interventions.

# Visualizations {data-navmenu="Visualizations"}

## Column {data-height="1000"}

### Obesity Prevalence Forest Plot

**Description:** This forest plot displays the **age-adjusted obesity prevalence** or **age-adjusted prevalence of no leisure physical activity** for counties in Georgia depending on your selection.\
- **Each dot** represents the estimated prevalence for a county.\
- **Horizontal lines** indicate confidence intervals, which show the range of possible values with statistical certainty.\
- Use the dropdown menu to select **counties** and compare.

**Key Takeaway:** The default forest plot shows the 10 counties with the highest age-adjusted prevalence of obesity in Georgia.

```{r, forest plot}
#ui component
ui <- fluidPage(
  titlePanel("Obesity and No Leisure Physical Activity Prevalence Across Georgia Counties"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("selected_counties", "Select Georgia Counties:", 
                  choices = unique(Obesity %>% filter(statedesc == "Georgia") %>% pull(locationname)), 
                  selected = NULL, multiple = TRUE),
      
      radioButtons("prevalence_type", "Choose Prevalence Type:",
                   choices = c("Obesity", "No Leisure Physical Activity"),
                   selected = "Obesity")
    ),
    
mainPanel(
      plotlyOutput("obesity_plot", height = "700px")
  )
 )
)


#server component
server <- function(input, output, session) {
  
  output$obesity_plot <- renderPlotly({
    req(input$prevalence_type)
    
    df <- if (input$prevalence_type == "Obesity") {
      Obesity
    } else {
      LPA
    }
    
    df <- df %>% filter(statedesc == "Georgia")
    
    state_avg <- df %>%
      summarise(mean_val = mean(data_value, na.rm = TRUE),
                lower_ci = mean(low_confidence_limit, na.rm = TRUE),
                upper_ci = mean(high_confidence_limit, na.rm = TRUE))
    
if (!is.null(input$selected_counties) && length(input$selected_counties) > 0) {
  county_data <- df %>% filter(locationname %in% input$selected_counties)
} else {
  county_data <- df %>%
    arrange(desc(data_value)) %>%
    slice_head(n = 10)
}

    county_data <- county_data %>% arrange(desc(data_value))
    plot_height <- max(500, nrow(county_data))
    ga_label <- paste("GA Average (", round(state_avg$mean_val, 1), "%)", sep = "")
    avg_line_data <- data.frame(x = state_avg$mean_val, label = ga_label)


# Forest Plot    
    p <- ggplot(county_data, aes(y = reorder(locationname, data_value), x = data_value)) +
      geom_point(aes(
        text = paste0(
          "County: ", locationname, "<br>",
          "Population: ", totalpopulation, "<br>",
          "Year: ", year, "<br>",
          input$prevalence_type, " Prevalence: ", round(data_value, 1), "%"
        )
      ), color = "#1f78b4", size = 1.5) +
      geom_errorbar(aes(xmin = low_confidence_limit, xmax = high_confidence_limit),
                    width = 0.2, color = "gray40") +
      geom_vline(data = avg_line_data,
           aes(xintercept = x, linetype = label, color = label),
           size = 0.8) +
scale_color_manual(name = NULL, values = setNames("red", ga_label)) +
scale_linetype_manual(name = NULL, values = setNames("dashed", ga_label)) +
      scale_x_continuous(limits = c(0, 100), labels = scales::percent_format(scale = 1)) +
      labs(title = paste(input$prevalence_type, "Prevalence\nin Selected Georgia Counties"),
           x = paste(input$prevalence_type, "Prevalence (%)"), y = "County") +
      theme_minimal() +
theme(
  plot.margin = margin(t = 30, r = 20, b = 20, l = 20),  # Add more top margin here
  axis.text.y = element_text(size = 8, hjust = 0),
  axis.title.y = element_text(margin = margin(r = 80))
)
    
    ggplotly(p, tooltip = "text") %>%
      layout(
        yaxis = list(automargin = TRUE, tickangle = 0),
        height = plot_height
      )
  })
}

shinyApp(ui = ui, server = server)
```

## Column {data-height="600"}
-----------------------------------------------------------------------

### Obesity and No Leisure Physical Activity Prevalence Heat Maps

**Description:** These interactive maps visualize county-level data for **obesity prevalence** and **prevalence of no leisure physical activity** in the state of Georgia.\
- Select a **county** from the dropdown to zoom into the county.\
- Use the **slider** to adjust the range of displayed prevalence values.\
- Click a county for more details.

**Key Takeaway:** The default heat maps show the prevalence of obesity and no leisure physical activity across the state of Georgia. It is evident that Macon and Randolph counties have a high prevalence of both **obesity** and **no leisure physical activity**.

```{r}

#ui component

ui_map <- fluidPage(
  titlePanel("Obesity and No Leisure Physical Activity Heat Maps"),
  
  fluidRow(
    column(6,
           wellPanel(
             selectInput("selected_map_county", "Choose a County in Georgia:", 
                         choices = c("All Counties", unique(Obesity %>% filter(statedesc == "Georgia") %>% pull(locationname))), 
                         selected = "All Counties"),
             sliderInput("prevalence_range", "Filter by Age-Adjusted Obesity Prevalence (%):",
                         min = 10, max = 53, value = c(10, 53), step = 1)
           ),
           leafletOutput("obesity_map", height = "800px")
    ),
    
    column(6,
           wellPanel(
             selectInput("selected_LPA_county", "Choose a County in Georgia:", 
                         choices = c("All Counties", unique(LPA %>% filter(statedesc == "Georgia") %>% pull(locationname))), 
                         selected = "All Counties"),
             sliderInput("prevalence_range_LPA", "Filter by Age-Adjusted Prevalence of No Leisure Physical Activity (%):",
                         min = 10, max = 50, value = c(10, 50), step = 1)
           ),
           leafletOutput("LPA_map", height = "800px")
    )
  )
)

#server component

server_map <- function(input, output, session) {
  
#Center of Georgia
get_georgia_center <- function(dataset) {
  state_data <- dataset %>% filter(statedesc == "Georgia")
  state_bbox <- st_bbox(state_data$geometry)
  lng_center <- mean(c(state_bbox["xmin"], state_bbox["xmax"]))
  lat_center <- mean(c(state_bbox["ymin"], state_bbox["ymax"]))
  return(list(lng = lng_center, lat = lat_center))
}
  
# **Obesity Map**
output$obesity_map <- renderLeaflet({
  # Filter only Georgia data
  map_data <- Obesity %>% filter(statedesc == "Georgia") %>%
    filter(data_value >= input$prevalence_range[1], 
           data_value <= input$prevalence_range[2])
  
  # Determine map center and zoom
  if (!is.null(input$selected_map_county) &&
      input$selected_map_county != "" &&
      input$selected_map_county != "All Counties") {
    
    selected_data <- map_data %>% filter(locationname == input$selected_map_county)
    if (nrow(selected_data) == 0) return(NULL)
    
    selected_geom <- selected_data$geometry[[1]]
    centroid <- st_centroid(selected_geom)
    lng_center <- st_coordinates(centroid)[1]
    lat_center <- st_coordinates(centroid)[2]
    zoom_level <- 9
    
    # Show only selected county
    map_data <- selected_data
    
  } else {
    georgia_center <- get_georgia_center(Obesity)
    lng_center <- georgia_center$lng
    lat_center <- georgia_center$lat
    zoom_level <- 7
  }
  
  if (nrow(map_data) == 0) return(NULL)
  
  pal <- colorNumeric(palette = "YlOrRd", domain = map_data$data_value)
  
  leaflet(map_data) %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    addPolygons(
      fillColor = ~pal(data_value),
      color = "black", weight = 0.5, opacity = 0.8,
      fillOpacity = 0.7,
      label = ~paste0(locationname, ": ", round(data_value, 1), "%"),
            popup = ~paste0("<b>State: </b>", statedesc,
                      "<br><b>County: </b>", locationname, 
                      "<br><b>Obesity Prevalence: </b>", round(data_value, 1), "%", 
                      "<br><b>95% CI: </b>[", low_confidence_limit, ", ", high_confidence_limit, "]")
    ) %>%
    addLegend(pal = pal, values = map_data$data_value, 
              title = "Obesity Prevalence (%)", position = "bottomright") %>%
    setView(lng = lng_center, lat = lat_center, zoom = zoom_level)
})

  
# **No Leisure Physical Activity Map**
output$LPA_map <- renderLeaflet({
  # Filter only Georgia data
  map_data <- LPA %>% filter(statedesc == "Georgia")
  
  # Prevalence range filter
  map_data <- map_data %>%
    filter(data_value >= input$prevalence_range_LPA[1], 
           data_value <= input$prevalence_range_LPA[2])
  
  # Map center and zoom
  if (!is.null(input$selected_LPA_county) &&
      input$selected_LPA_county != "" &&
      input$selected_LPA_county != "All Counties") {
    
    # Zoom selected county
    selected_data <- map_data %>% filter(locationname == input$selected_LPA_county)
    if (nrow(selected_data) == 0) return(NULL)
    
    selected_geom <- selected_data$geometry[[1]]
    centroid <- st_centroid(selected_geom)
    lng_center <- st_coordinates(centroid)[1]
    lat_center <- st_coordinates(centroid)[2]
    zoom_level <- 9
    
    # Just the selected county
    map_data <- selected_data
    
  } else {
    # Default: full Georgia view
    georgia_center <- get_georgia_center(LPA)
    lng_center <- georgia_center$lng
    lat_center <- georgia_center$lat
    zoom_level <- 7
  }
  
  if (nrow(map_data) == 0) return(NULL)
  
  # Color palette
  pal <- colorNumeric(palette = "Blues", domain = map_data$data_value)
  
  # Render map
  leaflet(map_data) %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    addPolygons(
      fillColor = ~pal(data_value),
      color = "black", weight = 0.5, opacity = 0.8,
      fillOpacity = 0.7,
      label = ~paste0(locationname, ": ", round(data_value, 1), "%"),
      popup = ~paste0("<b>State: </b>", statedesc,
                      "<br><b>County: </b>", locationname, 
                      "<br><b>No Leisure Physical Activity: </b>", round(data_value, 1), "%",
                      "<br><b>95% CI: </b>[", low_confidence_limit, ", ", high_confidence_limit, "]")
    ) %>%
    addLegend(pal = pal, values = map_data$data_value, 
              title = "No Leisure Physical Activity (%)", position = "bottomright") %>%
    setView(lng = lng_center, lat = lat_center, zoom = zoom_level)
})
}

shinyApp(ui = ui_map, server = server_map)

```


```{r}
rsconnect::setAccountInfo(name='stephanie-a-ponce',
			  token='5EC29BE421999A79014DB3305315AD89',
			  secret='22ZDfpvDvhDzqsKpIxTPUXS3unsxPU5vMLIXeu9o')
```
